### 2023-04-06 extension.rmd
Add post-acceptance edits by Adrian

---
title             : "Online appendix to \"Bayesian multilevel modeling and its application in comparative communication research\""
shorttitle        : "BAYESIAN MULTILEVEL REGRESSION"

@@

```{r clplot, echo = FALSE, fig.cap = "Posterior predictive checks with 100 sets of simulated data based on the multivariate ordinal model"}
require(cowplot)
c9c10ppc <- readRDS("ppa_clmod_ppc.RDS")
c9c10ppc[[1]] + ggtitle(label = "", subtitle = "How much freedom do you personally have in selecting news stories you work on?") -> c9gg
c9c10ppc[[2]] + ggtitle(label = "", subtitle = "How much freedom do you personally have in deciding which aspects of a story should \n be emphasized?") -> c10gg
plot_grid(c9gg, c10gg, ncol = 1)
```

>>>>

---
title             : "Online appendix to \"Bayesian multilevel modeling and its application in comparative journalism studies\""
shorttitle        : "BAYESIAN MULTILEVEL REGRESSION"

@@

```{r clplot, echo = FALSE, fig.cap = "Posterior predictive checks with 100 sets of simulated data based on the multivariate ordinal model"}
require(cowplot)
require(ggplot2)
c9c10ppc <- readRDS("ppa_clmod_ppc.RDS")
c9c10ppc[[1]] + ggtitle(label = "", subtitle = "How much freedom do you personally have in selecting news stories you work on?") -> c9gg
c9c10ppc[[2]] + ggtitle(label = "", subtitle = "How much freedom do you personally have in deciding which aspects of a story should \n be emphasized?") -> c10gg
plot_grid(c9gg, c10gg, ncol = 1)
```

### 2023-04-06 manuscript.rmd
Add post-acceptance edits by Adrian

author: 
  - name          : "Chung-hong Chan"
    affiliation   : "1"
    corresponding : yes
    address       : "Unter Sachsenhausen 6-8, 50667 Köln"
    email         : "chung-hong.chan@gesis.org"
  - name          : "Adrian Rauschfleisch"
    affiliation   : "2"

@@

authornote: |
  The authors would like to thank professor Kasper Welbers (Vrije Universiteit Amsterdam) for his advice on the development of the Dutch dictionary; Dr Junior Yuner Zhu (Hong Kong Baptist University) for sharing details about her study. The authors report there are no competing interests to declare.
  Source code and data are available at (redacted). Prereg: https://osf.io/2h4w8/

@@

Although some comparative studies are purely descriptive, most of these studies are aimed to study how contextual conditions such as characteristics of media outlets (e.g. political orientation, online versus offline) or their countries (e.g. post-communist country) influence the behaviors of journalists or the content of news articles. For example, some hypotheses are "the coverage of foreign countries by the news is primarily determined by geographical proximity and the status of the covered country" [@wilke:2012, p.306] and "the degree of opinion-orientation will be highest in newspapers from Polarized Mediterranean systems and lowest in those from Anglo-American systems." [@esser:2013:C] Employing the language from epidemiology [@susser:1994], the effect on media content and journalistic behaviors in these hypotheses is assumed to be ecological. It assumes a macro contextual factor (e.g., newspapers from Polarized Mediterranean systems) is associated with an outcome at the micro-level (e.g., degree of opinion-orientation), namely, the journalist-level or article-level.

This study of ecological effects on individual behaviors has a long tradition in communication research. There are hypothesis-generating and hypothesis-testing comparative studies. An example of hypothesis-generating comparative research is <!-- redacted for ijoc  -->AUTHOR1 <!-- @chan2020combining -->, in which the sentiment profiles of terrorism coverage from Muslim- and Christian-majority countries were visualized. Hypothesis-generating comparative study, however, is rare. As we see from the example hypotheses above, most of these studies propose hypotheses to test for ecological effects. In this context usually traditional (frequentist) hypothesis testing approaches were used: @esser:2013:C lumped multiple outlets from the same country together and then used univariate analysis of variance to test for the differences in the proportion of opinion-orienting articles across countries. In @wu:2000:SDI, multiple stepwise regression was used. These approaches violate the underlying independence of observations assumption of linear regression (See non-atomicity below). It highlights the fact that comparative research introduces a feature that researchers usually overlook: media contents are clustered in a multilevel structure. A news article is nested within its media outlet and its media outlet is, in turn, nested within its country. Such data structure brings two problems: non-atomicity and non-stochasticity. And specifically, we propose two solutions: multilevel modeling and Bayesian approach.

The goal of this paper is twofold. We use examples from journalism studies as the most critical cases in communication science to highlight the strength of multilevel models and the advantages of Bayesian models. In our paper, we first discuss these two issues and then suggest Bayesian multilevel regression as the most defensible option. Using the openly available WJS and *useNews* datasets, we then demonstrate how to apply a Bayesian approach to analyze comparative data.

@@

This method is the so-called "disaggregation approach". From an epistemic standpoint, searching for ecological effect by studying the value of slope ($\beta_{1}$) leads to the so-called atomistic fallacy [@hox2017multilevel]. From a statistical standpoint, this approach violates the underlying independence assumption. Almost all frequentist statistical tests assume that the observations are independent of each other: $y_{i}$ given $x_{i}$ are independently identically distributed (i.i.d.). Journalists from the same country are hardly i.i.d.: they are subjected to being governed by the same government and exposed to the same journalistic culture. Also, the way communication researchers survey the journalists resembles (nonrandom) cluster sampling rather than pure random sampling: one usually starts from a specific country and then surveys journalists from it. The choice of countries is almost always not randomly selected. This sampling approach makes the independence assumption even more fragile. Previous simulation studies have shown that ignoring this dependence can lead to false-positive associations [e.g., @clarke:2008:W;@chen:2012:IIL].

@@

Another solution is to use the multilevel model (linear mixed model, or hierarchical model). In a multilevel model, the effect on the micro-level dependent variable ($y$) is modeled with equations at different levels. Using the above example, $y_{ik}$ denotes the perceived professional autonomy of the $i$-th journalist in the $j$-th country; $x_{k}$ denotes the democratic performance of the $k$-th country.

@@

In these equations, $\gamma_{00}$ is the average slope, while $\mu_{0k}$ is group-dependent deviations of the slope from the average. It is usually set as having a normal distribution with a variance $\tau_{00}$, i.e. $\mu_{0k} \sim \mathcal{N}(0, \tau_{00})$. Instead of a single value, the regression coefficient $\gamma_{00}$ is assumed to be a distribution of values depending on a macro-level group [^pooling]. It addresses the problem of clustering of articles by macro-level variables. This model is called the varying-intercept model and is used frequently in social science research. We can then study the magnitude of $\gamma_{01}$ to determine the ecological effect.

[^pooling]: One can also think about this in terms of pooling: how the data from different groups are pooled together. Complete pooling (the disaggregation approach, one slope for all groups), no pooling (each group is modelled individually and each group has a different slope), and partial pooling (each group has a unique slope. But the distribution of all slopes follows a distribution) are three scenarios. Multilevel modeling uses partial pooling.

@@

The Bayesian multilevel regression gave the *posterior distribution*, $P(\theta | X)$, of the regression coefficients. It is a distribution and therefore we need ways to display both the central tendency and spread of the distribution. By default, *brms* displays the mean and the 95\% High Density Interval (HDI). <!-- [^5] --> These two values represent the point and interval estimates of the regression coefficient.

@@

```{r}
wjs_reg <- readRDS("wjs_reg.RDS")
wjs_reg$term <- c("(Intercept)", "Experience", "Rank", "Gender (Female)", "University Degree", "Index of Democracy", "GDP per capita")
colnames(wjs_reg)[2:4] <- c("Bayesian", "MLE", "Disaggregation")
apa_table(wjs_reg, caption = "Point and interval estimates from WJS analysis using three different analytic strategies: Bayesian multilevel modeling, multilevel modeling using MLE, and Disaggregation approach")
```

@@

@@

\begin{align}
\log{z_{jk}} &= \gamma_{00} + \mu_{0k} + \gamma_{01} \log{x_{k}} + \log{n_j} + \epsilon_{jk} \\
\log{z_{jk}} - \log{n_j} &= \gamma_{00} + \mu_{0k} + \gamma_{01} \log{x_{k}} +  \epsilon_{jk} \\
\log\frac{z_{jk}}{n_{j}} &= \gamma_{00} + \mu_{0k} + \gamma_{01} \log{x_{k}} + \epsilon_{jk}\end{align}

@@

```{r}
import_reg <- readRDS("import_reg.RDS")
import_reg$term <- c("(Intercept)", "Log (Trade Volume)")
colnames(import_reg)[2:4] <- c("Bayesian", "MLE", "Disaggregation")
apa_table(import_reg, caption = "Point and interval estimates from outlet-level analysis using three different analytic strategies: Bayesian multilevel modeling, multilevel modeling using MLE, and Disaggregation approach")
```

@@

Similar to the previous example, we conducted the posterior predictive check (Figure  \@ref(fig:fig3x)). Our model can capture both the range, peak and shape of the original data. But there are many variations in the simulated data, probably due to the fact that the Bayesian model is based on just 61 data points and one single predictor.

@@

In this paper, we argue the case for using Bayesian multilevel regression analysis to analyze data from comparative communication research. We demonstrate using the openly available WJS and useNews datasets that Bayesian analysis provides valid inference of ecological effects and can be done easily with the R package *brms*. We mainly used the WJS dataset to illustrate the strength of multilevel models, whereas the useNews dataset was selected as an example where Bayesian models are conceptually the most defensible option. However, both multilevel models and Bayesian analysis are connected approaches.

@@

Using a Bayesian approach still has several limitations. First, it takes more time to get the results of a Bayesian regression analysis. Estimating a Bayesian regression model is a computationally demanding task. It took four days on a regular computer to get the results of the Bayesian regression model with 1,525,871 articles (Example 2, article-level analysis). Still, eventually, it worked, and typical studies in comparative journalism research , e.g. 27,567 journalists in Example 1 can be estimated within a reasonable time frame. Secondly, if sampling procedures are used for a typical experimental design that could be replicated, using a frequentist approach is probably the less complicated approach. In any other scenario, we believe the benefits of the Bayesian approach outweigh the potential limitations.

>>>>

author: 
  - name          : "Chung-hong Chan"
    affiliation   : "1"
    corresponding : yes
    address       : "Unter Sachsenhausen 6-8, 50667 Köln"
    email         : "chung-hong.chan@gesis.org"
  - name          : "Adrian Rauchfleisch"
    affiliation   : "2"

@@

authornote: |
  Adrian Rauchfleisch’s work was supported by the National Science and Technology Council, Taiwan (ROC), grant no. 111-2628-H-002-003-. The authors would like to thank professor Kasper Welbers (Vrije Universiteit Amsterdam) for his advice on the development of the Dutch dictionary; Professor Junior Yuner Zhu (Hong Kong Baptist University) for sharing details about her study.
  Source code and data are available at https://github.com/chainsawriot/usenews_bayes. Preregistration: https://osf.io/2h4w8/

@@

Although some comparative studies are purely descriptive, most of these studies are aimed to study how contextual conditions such as characteristics of media outlets (e.g. political orientation, online versus offline) or their countries (e.g. post-communist country) influence journalists' behaviors or the content of news articles. For example, some hypotheses are "the coverage of foreign countries by the news is primarily determined by geographical proximity and the status of the covered country" [@wilke:2012, p.306] and "the degree of opinion-orientation will be highest in newspapers from Polarized Mediterranean systems and lowest in those from Anglo-American systems." [@esser:2013:C] Employing the language from epidemiology [@susser:1994], the effect on media content and journalistic behaviors in these hypotheses is assumed to be ecological. It assumes a macro contextual factor (e.g., newspapers from Polarized Mediterranean systems) is associated with an outcome at the micro-level (e.g., degree of opinion-orientation), namely, the journalist-level or article-level.

This study of ecological effects on individual behaviors has a long tradition in communication research. There are hypothesis-generating and hypothesis-testing comparative studies. An example of hypothesis-generating comparative research is @chan2020combining, in which the sentiment profiles of terrorism coverage from Muslim- and Christian-majority countries were visualized. Hypothesis-generating comparative study, however, is rare. As we see from the example hypotheses above, most of these studies propose hypotheses to test for ecological effects. Usually traditional (frequentist) hypothesis testing approaches were used in this context: @esser:2013:C lumped multiple outlets from the same country together and then used univariate analysis of variance to test for the differences in the proportion of opinion-orienting articles across countries. In @wu:2000:SDI, multiple stepwise regression was used. These approaches violate the underlying independence of observations assumption of linear regression (See non-atomicity below). It highlights the fact that comparative research introduces a feature that researchers usually overlook: media contents are clustered in a multilevel structure. A news article is nested within its media outlet and its media outlet is, in turn, nested within its country. Such data structure brings two problems: non-atomicity and non-stochasticity. And specifically, we propose two solutions: multilevel modeling and Bayesian approach.

The goal of this paper is twofold. We use examples from journalism studies as the most critical cases in communication science to highlight (1) the strength of multilevel models and (2) the advantages of Bayesian models. In our paper, we first discuss these two issues and then suggest Bayesian multilevel regression as the most defensible option. Using the openly available WJS and *useNews* datasets, we then demonstrate how to apply a Bayesian approach to analyze comparative data.

@@

This method is the so-called "disaggregation approach". From an epistemic standpoint, searching for an ecological effect by studying the value of slope ($\beta_{1}$) leads to the so-called atomistic fallacy [@hox2017multilevel]. From a statistical standpoint, this approach violates the underlying independence assumption. Almost all frequentist statistical tests assume that the observations are independent of each other: $y_{i}$ given $x_{i}$ are independently identically distributed (i.i.d.). Journalists from the same country are hardly i.i.d.: they are subjected to being governed by the same government and exposed to the same journalistic culture. Also, the way communication researchers survey the journalists resembles (nonrandom) cluster sampling rather than pure random sampling: one usually starts from a specific country and then surveys journalists from it. The choice of countries is almost always not randomly selected. This sampling approach makes the independence assumption even more fragile. Previous simulation studies have shown that ignoring this dependence can lead to false-positive associations [e.g., @clarke:2008:W;@chen:2012:IIL].

@@

Another solution is to use the multilevel model (linear mixed model, or hierarchical model). In a multilevel model, the effect on the micro-level dependent variable ($y$) is modeled with equations at different levels. Using the above example, $y_{ik}$ denotes the perceived professional autonomy of the $i$-th journalist in the $k$-th country; $x_{k}$ denotes the democratic performance of the $k$-th country.

@@

In these equations, $\gamma_{00}$ is the average slope, while $\mu_{0k}$ is group-dependent deviations of the slope from the average. It is usually set as having a normal distribution with a variance $\tau_{00}$, i.e. $\mu_{0k} \sim \mathcal{N}(0, \tau_{00})$. Instead of a single value, the regression coefficient $\gamma_{00}$ is assumed to be a distribution of values depending on a macro-level group [^pooling]. It addresses the problem of clustering of articles by macro-level variables. This model is called the varying-intercept model and is frequently used in social science research. We can then study the magnitude of $\gamma_{01}$ to determine the ecological effect.

[^pooling]: One can also think about this in terms of pooling: how the data from different groups are pooled together. Complete pooling (the disaggregation approach, one slope), no pooling (each group is modelled individually and each group has a different slope), and partial pooling (each group has a unique slope. But the distribution of all slopes follows a distribution) are three scenarios. Multilevel modeling uses partial pooling.

@@

The Bayesian multilevel regression gave the *posterior distribution*, $P(\theta | X)$, of the regression coefficients. It is a distribution and therefore we need ways to display both the central tendency and spread of the distribution. By default, *brms* displays the mean and the 95\% Highest Density Interval (HDI). <!-- [^5] --> These two values represent the point and interval estimates of the regression coefficient.

@@

```{r}
wjs_reg <- readRDS("wjs_reg.RDS")
wjs_reg$term <- c("(Intercept)", "Experience", "Rank", "Gender (Female)", "University Degree", "Index of Democracy", "GDP per capita")
colnames(wjs_reg)[2:4] <- c("Bayesian", "MLE", "Disaggregation")
apa_table(wjs_reg, caption = "Point and Interval Estimates From WJS Analysis Using Three Different Analytic Strategies: Bayesian Multilevel Modeling, Multilevel Modeling Using MLE, and Disaggregation Approach")
```

@@

\begin{align}
\log{z_{jk}} &= \gamma_{00} + \mu_{0k} + \gamma_{01} \log{x_{k}} + \log{n_j} + \epsilon_{jk} \\
\log{z_{jk}} - \log{n_j} &= \gamma_{00} + \mu_{0k} + \gamma_{01} \log{x_{k}} +  \epsilon_{jk} \\
\log\frac{z_{jk}}{n_{j}} &= \gamma_{00} + \mu_{0k} + \gamma_{01} \log{x_{k}} + \epsilon_{jk}
\end{align}

@@

@@

```{r}
import_reg <- readRDS("import_reg.RDS")
import_reg$term <- c("(Intercept)", "Log (Trade Volume)")
colnames(import_reg)[2:4] <- c("Bayesian", "MLE", "Disaggregation")
apa_table(import_reg, caption = "Point and Interval Estimates From Outlet-Level Analysis Using Three Different Analytic Strategies: Bayesian Multilevel Modeling, Multilevel Modeling Using MLE, and Disaggregation Approach")
```

@@

Similar to the previous example, we conducted the posterior predictive check (Figure  \@ref(fig:fig3x)). Our model can capture the original data's range, peak, and shape. But there are many variations in the simulated data, probably due to the fact that the Bayesian model is based on just 61 data points and one single predictor.

@@

In this paper, we argue the case for using Bayesian multilevel regression analysis to analyze data from comparative communication research. Using the openly available WJS and useNews datasets, we demonstrate that Bayesian analysis provides valid inference of ecological effects and can be done easily with the R package *brms*. We mainly used the WJS dataset to illustrate the strength of multilevel models, whereas the useNews dataset was selected as an example where Bayesian models are conceptually the most defensible option. However, both multilevel models and Bayesian analysis are connected approaches.

@@

Using a Bayesian approach still has several limitations. First, it takes more time to get the results of a Bayesian regression analysis. Estimating a Bayesian regression model is a computationally demanding task. It took four days on a regular computer to get the results of the Bayesian regression model with 1,525,871 articles (Example 2, article-level analysis). Still, eventually, it worked, and typical studies in comparative journalism research , e.g. 27,567 journalists in Example 1 can be estimated within a reasonable time frame (< 10 mins). Secondly, if sampling procedures are used for a typical experimental design that could be replicated, using a frequentist approach is probably the less complicated approach. In any other scenario, we believe the benefits of the Bayesian approach outweigh the potential limitations.
